Inference methods comparison for the simplified Naomi model using the Kolmogorov-Smirnov test statistic

### 2023-07-03 src/naomi-simple_ks/ks.Rmd
Simplify KS comparison, ready to use AD as well if required

write_csv(time_taken, "time_taken.csv")

@@

```{r}
histogram_and_ecdf_helper <- function(par) lapply(1:sum(names(tmb$fit$obj$env$par) == par), histogram_and_ecdf, par = par)
```

@@

```{r eval = pars_eval[["beta_rho"]]}
histogram_and_ecdf_helper("beta_rho")
```

@@

```{r eval = pars_eval[["beta_alpha"]]}
histogram_and_ecdf_helper("beta_alpha")
```

@@

```{r eval = pars_eval[["beta_lambda"]]}
histogram_and_ecdf_helper("beta_lambda")
```

@@

```{r eval = pars_eval[["beta_anc_rho"]]}
histogram_and_ecdf_helper("beta_anc_rho")
```

@@

```{r eval = pars_eval[["u_rho_x"]]}
histogram_and_ecdf_helper("u_rho_x")
```

@@

```{r eval = pars_eval[["us_rho_x"]]}
histogram_and_ecdf_helper("us_rho_x")
```

@@

```{r eval = pars_eval[["u_rho_xs"]]}
histogram_and_ecdf_helper("u_rho_xs")
```

@@

```{r eval = pars_eval[["us_rho_xs"]]}
histogram_and_ecdf_helper("us_rho_xs")
```

@@

```{r eval = pars_eval[["u_rho_as"]]}
histogram_and_ecdf_helper("u_rho_as")
```

@@

```{r eval = pars_eval[["u_alpha_x"]]}
histogram_and_ecdf_helper("u_alpha_x")
```

@@

```{r eval = pars_eval[["us_alpha_x"]]}
histogram_and_ecdf_helper("us_alpha_x")
```

@@

```{r eval = pars_eval[["u_alpha_xs"]]}
histogram_and_ecdf_helper("u_alpha_xs")
```

@@

```{r eval = pars_eval[["beta_rho"]]}
histogram_and_ecdf_helper("us_alpha_xs")
```

@@

```{r eval = pars_eval[["u_alpha_a"]]}
histogram_and_ecdf_helper("u_alpha_a")
```

@@

```{r eval = pars_eval[["u_alpha_as"]]}
histogram_and_ecdf_helper("u_alpha_as")
```

@@

```{r eval = pars_eval[["u_alpha_xa"]]}
histogram_and_ecdf_helper("u_alpha_xa")
```

@@

```{r eval = pars_eval[["ui_lambda_x"]]}
histogram_and_ecdf_helper("ui_lambda_x")
```

@@

```{r eval = pars_eval[["ui_anc_rho_x"]]}
histogram_and_ecdf_helper("ui_anc_rho_x")
```

@@

```{r eval = pars_eval[["ui_anc_alpha_x"]]}
histogram_and_ecdf_helper("ui_anc_alpha_x")
```

@@

```{r eval = pars_eval[["log_or_gamma"]]}
histogram_and_ecdf_helper("log_or_gamma")
```

@@

@@

```{r}
ks_helper <- function(par, starts_with = FALSE, ...) {
  to_ks_df(par, starts_with = starts_with) %>% ks_plot(par, ...)
}

@@

```{r}
ks_helper("beta", starts_with = TRUE)
```

@@

```{r}
ks_helper("logit", starts_with = TRUE)
```

@@

```{r}
ks_helper("log_sigma", starts_with = TRUE)
```

@@

```{r eval = pars_eval[["u_rho_x"]]}
ks_helper("u_rho_x")
```

@@

```{r eval = pars_eval[["u_rho_xs"]]}
ks_helper("u_rho_xs")
```

@@

```{r eval = pars_eval[["us_rho_x"]]}
ks_helper("us_rho_x")
```

@@

```{r eval = pars_eval[["us_rho_xs"]]}
ks_helper("us_rho_xs")
```

@@

```{r eval = pars_eval[["u_rho_a"]]}
ks_helper("u_rho_a")
```

@@

```{r eval = pars_eval[["u_rho_as"]]}
ks_helper("u_rho_as")
```

@@

```{r eval = pars_eval[["u_alpha_x"]]}
ks_helper("u_alpha_x")
```

@@

```{r eval = pars_eval[["u_alpha_xs"]]}
ks_helper("u_alpha_xs")
```

@@

```{r eval = pars_eval[["us_alpha_x"]]}
ks_helper("us_alpha_x")
```

@@

```{r eval = pars_eval[["us_alpha_xs"]]}
ks_helper("us_alpha_xs")
```

@@

```{r eval = pars_eval[["u_alpha_a"]]}
ks_helper("u_alpha_a")
```

@@

```{r eval = pars_eval[["u_alpha_as"]]}
ks_helper("u_alpha_as")
```

@@

```{r eval = pars_eval[["u_alpha_xa"]]}
ks_helper("u_alpha_xa")
```

@@

```{r eval = pars_eval[["ui_anc_rho_x"]]}
ks_helper("ui_anc_rho_x")
```

@@

```{r eval = pars_eval[["ui_anc_alpha_x"]]}
ks_helper("ui_anc_alpha_x")
```

@@

```{r eval = pars_eval[["log_or_gamma"]]}
ks_helper("log_or_gamma")
```

@@

```{r}
# ks_lambda_t1_out %>%
#   filter(ks == 0) %>%
#   group_by(method) %>%
#   summarise(n = n())
# 
# ks_lambda_t1_out %>%
#   filter(ks == 0) %>%
#   pull(age_group) %>%
#   unique()
```

@@

Values in dark blue are the lowest KS difference for that particular parameter:

@@

ks_summary <- lapply(unique(names(tmb$fit$obj$env$par)), function(x) {
  to_ks_df(x) %>%
    group_by(method, index) %>%
    summarise(ks = mean(ks), par = x) %>%
    ungroup()
}) %>%
  bind_rows() %>%
  pivot_wider(names_from = "method", values_from = "ks") %>%
  rename(
    "Parameter" = "par",
    "KS(aghq, tmbstan)" = "aghq",
    "KS(TMB, tmbstan)" = "TMB",
  )

r <- tmb$obj$env$random
x_names <- names(tmb$obj$env$par[r])
theta_names <- names(tmb$obj$env$par[-r])

dict <- data.frame(
  Parameter = c(unique(x_names), unique(theta_names)),
  Type = c(rep("Latent field", length(unique(x_names))), rep("Hyper", length(unique(theta_names))))
)

ks_summary <- ks_summary %>%
  left_join(dict, by = "Parameter")

saveRDS(ks_summary, "ks_summary.rds")

ks_summary %>%
  filter(Type != "Hyper") %>%
  ks_plot_many("TMB", "aghq")
```

```{r}
ks_summary %>%
  group_by(Type) %>%
  summarise(
    TMB = mean(`KS(TMB, tmbstan)`),
    aghq = mean(`KS(aghq, tmbstan)`)
  ) %>%
  gt::gt()
```

@@

```{r}
ks_summary %>%
  filter(Type == "Latent field") %>%
  group_by(Parameter) %>%
  summarise(
    `KS(TMB, tmbstan)` = mean(`KS(TMB, tmbstan)`),
    `KS(aghq, tmbstan)` = mean(`KS(aghq, tmbstan)`)
  ) %>%
  ks_table()
```

### Hyperparameters

```{r}
# ks_summary %>%
#   filter(Type == "Hyper") %>%
#   group_by(Parameter) %>%
#   summarise(
#     `KS(TMB, tmbstan)` = mean(`KS(TMB, tmbstan)`),
#     `KS(aghq, tmbstan)` = mean(`KS(aghq, tmbstan)`)
#   ) %>%
#   ks_table()
```

### Outputs

```{r}
#' To-do
```

@@

```{r}
ks_comparison <- function(ks_summary, method1, method2, n) {
  ks_method1 <- paste0("KS(", method1, ", tmbstan)")
  ks_method2 <- paste0("KS(", method2, ", tmbstan)")
  ks_summary[["KS difference"]] <- ks_summary[[ks_method1]] - ks_summary[[ks_method2]]
  ks_summary %>%
    filter(Type != "Hyper") %>%
    arrange(desc(`KS difference`)) %>%
    head(n = n)
}
```

@@

```{r}
ks_comparison(ks_summary, method1 = "aghq", method2 = "TMB", 10)

histogram_and_ecdf(par = "us_rho_x", i = 19)
histogram_and_ecdf(par = "ui_anc_rho_x", i = 2)
histogram_and_ecdf(par = "u_rho_x", i = 2)
```

@@

```{r}
ks_comparison(ks_summary, method1 = "TMB", method2 = "aghq", 10)

histogram_and_ecdf(par = "us_alpha_xs", i = 18)
histogram_and_ecdf(par = "us_alpha_xs", i = 30)
histogram_and_ecdf(par = "u_alpha_xa", i = 10)
```

@@

ks_summary %>%
  pivot_longer(
    col = starts_with("KS"),
    names_to = "Method",
    values_to = "KS"
  ) %>%
  mutate(
    Method = fct_recode(Method,
      "aghq" = "KS(aghq, tmbstan)",
      "TMB" = "KS(TMB, tmbstan)"
    )
  ) %>%
  group_by(Parameter) %>%
  mutate(
    par_num = case_when(
      max(index) > 1 ~ paste0(Parameter, "[", index, "]"),
      TRUE ~ Parameter
    )
  ) %>%
  ungroup() %>%
  left_join(data.frame(ess) %>%
    tibble::rownames_to_column("par_num")
  ) %>%
  ggplot(aes(x = ess, y = KS)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", color = "#56B4E9") +
    facet_grid(~ Method) +
    theme_minimal() +
    labs(x = "ESS", y = "KS(method, tmbstan)")
```

>>>>

write_csv(time_taken, "time-taken.csv")

@@

@@

```{r eval = pars_eval[["beta_rho"]]}
histogram_and_ecdf_list("beta_rho")
```

@@

```{r eval = pars_eval[["beta_alpha"]]}
histogram_and_ecdf_list("beta_alpha")
```

@@

```{r eval = pars_eval[["beta_lambda"]]}
histogram_and_ecdf_list("beta_lambda")
```

@@

```{r eval = pars_eval[["beta_anc_rho"]]}
histogram_and_ecdf_list("beta_anc_rho")
```

@@

```{r eval = pars_eval[["u_rho_x"]]}
histogram_and_ecdf_list("u_rho_x")
```

@@

```{r eval = pars_eval[["us_rho_x"]]}
histogram_and_ecdf_list("us_rho_x")
```

@@

```{r eval = pars_eval[["u_rho_xs"]]}
histogram_and_ecdf_list("u_rho_xs")
```

@@

```{r eval = pars_eval[["us_rho_xs"]]}
histogram_and_ecdf_list("us_rho_xs")
```

@@

```{r eval = pars_eval[["u_rho_as"]]}
histogram_and_ecdf_list("u_rho_as")
```

@@

```{r eval = pars_eval[["u_alpha_x"]]}
histogram_and_ecdf_list("u_alpha_x")
```

@@

```{r eval = pars_eval[["us_alpha_x"]]}
histogram_and_ecdf_list("us_alpha_x")
```

@@

```{r eval = pars_eval[["u_alpha_xs"]]}
histogram_and_ecdf_list("u_alpha_xs")
```

@@

```{r eval = pars_eval[["beta_rho"]]}
histogram_and_ecdf_list("us_alpha_xs")
```

@@

```{r eval = pars_eval[["u_alpha_a"]]}
histogram_and_ecdf_list("u_alpha_a")
```

@@

```{r eval = pars_eval[["u_alpha_as"]]}
histogram_and_ecdf_list("u_alpha_as")
```

@@

```{r eval = pars_eval[["u_alpha_xa"]]}
histogram_and_ecdf_list("u_alpha_xa")
```

@@

```{r eval = pars_eval[["ui_lambda_x"]]}
histogram_and_ecdf_list("ui_lambda_x")
```

@@

```{r eval = pars_eval[["ui_anc_rho_x"]]}
histogram_and_ecdf_list("ui_anc_rho_x")
```

@@

```{r eval = pars_eval[["ui_anc_alpha_x"]]}
histogram_and_ecdf_list("ui_anc_alpha_x")
```

@@

```{r eval = pars_eval[["log_or_gamma"]]}
histogram_and_ecdf_list("log_or_gamma")
```

@@

```{r}
r <- tmb$fit$obj$env$random
x_names <- names(tmb$fit$obj$env$par[r])
theta_names <- names(tmb$fit$obj$env$par[-r])

dict <- data.frame(
  parname = c(unique(x_names), unique(theta_names)),
  type = c(rep("Latent field", length(unique(x_names))), rep("Hyper", length(unique(theta_names))))
)

ks_df <- lapply(unique(names(tmb$fit$obj$env$par)), to_ks_df) %>%
  bind_rows() %>%
  mutate(parname = str_extract(par, ".*(?=\\[)")) %>%
  left_join(dict, by = "parname")
```

@@

```{r}
ks_plot_starts_with <- function(param) {
  ks_df %>%
    filter(startsWith(par, param)) %>%
    ks_plot(par = param)
}

@@

```{r}
ks_plot_starts_with("beta")
```

@@

```{r}
ks_plot_starts_with("logit")
```

@@

```{r}
ks_plot_starts_with("log_sigma")
```

@@

```{r eval = pars_eval[["u_rho_x"]]}
ks_plot_starts_with("u_rho_x")
```

@@

```{r eval = pars_eval[["u_rho_xs"]]}
ks_plot_starts_with("u_rho_xs")
```

@@

```{r eval = pars_eval[["us_rho_x"]]}
ks_plot_starts_with("us_rho_x")
```

@@

```{r eval = pars_eval[["us_rho_xs"]]}
ks_plot_starts_with("us_rho_xs")
```

@@

```{r eval = pars_eval[["u_rho_a"]]}
ks_plot_starts_with("u_rho_a")
```

@@

```{r eval = pars_eval[["u_rho_as"]]}
ks_plot_starts_with("u_rho_as")
```

@@

```{r eval = pars_eval[["u_alpha_x"]]}
ks_plot_starts_with("u_alpha_x")
```

@@

```{r eval = pars_eval[["u_alpha_xs"]]}
ks_plot_starts_with("u_alpha_xs")
```

@@

```{r eval = pars_eval[["us_alpha_x"]]}
ks_plot_starts_with("us_alpha_x")
```

@@

```{r eval = pars_eval[["us_alpha_xs"]]}
ks_plot_starts_with("us_alpha_xs")
```

@@

```{r eval = pars_eval[["u_alpha_a"]]}
ks_plot_starts_with("u_alpha_a")
```

@@

```{r eval = pars_eval[["u_alpha_as"]]}
ks_plot_starts_with("u_alpha_as")
```

@@

```{r eval = pars_eval[["u_alpha_xa"]]}
ks_plot_starts_with("u_alpha_xa")
```

@@

```{r eval = pars_eval[["ui_anc_rho_x"]]}
ks_plot_starts_with("ui_anc_rho_x")
```

@@

```{r eval = pars_eval[["ui_anc_alpha_x"]]}
ks_plot_starts_with("ui_anc_alpha_x")
```

@@

```{r eval = pars_eval[["log_or_gamma"]]}
ks_plot_starts_with("log_or_gamma")
```

@@

@@

@@

ks_summary <- ks_df %>%
  group_by(method, parname, type) %>%
  summarise(ks = signif(mean(ks), 4)) %>%
  pivot_wider(names_from = "method", values_from = "ks") %>%
  rename(
    "Parameter" = "parname",
    "KS(aghq, tmbstan)" = "aghq",
    "KS(TMB, tmbstan)" = "TMB",
  )

saveRDS(ks_summary, "ks-summary.rds")

ks_summary %>%
  filter(type != "Hyper") %>%
  ks_plot_many("TMB", "aghq")
```

```{r}
ks_summary %>%
  group_by(type) %>%
  summarise(
    TMB = mean(`KS(TMB, tmbstan)`),
    aghq = mean(`KS(aghq, tmbstan)`)
  ) %>%
  gt::gt()
```

@@

```{r}
ks_summary %>%
  filter(type == "Latent field") %>%
  mutate(diff = signif(`KS(TMB, tmbstan)` - `KS(aghq, tmbstan)`, 3)) %>%
  DT::datatable()
```

@@

```{r}
ks_df_wide <- ks_df %>%
  pivot_wider(names_from = "method", values_from = "ks") %>%
  mutate(diff = TMB - aghq)
```

@@

```{r}
ks_df_wide %>%
  arrange(desc(diff)) %>%
  head(n = 10)

histogram_and_ecdf(par = "us_alpha_xs", i = 18)
histogram_and_ecdf(par = "us_alpha_xs", i = 16)
histogram_and_ecdf(par = "u_alpha_xs", i = 13)
```

@@

```{r}
ks_df_wide %>%
  arrange(diff) %>%
  head(n = 10)

histogram_and_ecdf(par = "ui_anc_rho_x", i = 27)
histogram_and_ecdf(par = "ui_anc_alpha_x", i = 19)
histogram_and_ecdf(par = "u_alpha_x", i = 18)
```

@@

ks_df %>%
  filter(type != "Hyper") %>%
  left_join(data.frame(ess) %>%
    tibble::rownames_to_column("par"),
  ) %>%
  ggplot(aes(x = ess, y = ks)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", color = "#56B4E9") +
    facet_grid(~ method) +
    theme_minimal() +
    labs(x = "ESS", y = "KS(method, tmbstan)")
```
